Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-07-18 edubots_surveys.Rnw
adds subcaptions and splits up figures refs #47

%%% Load packages
\usepackage{amsthm,amsmath}
\usepackage{multirow}
\usepackage{rotating}

\usepackage[utf8]{inputenc} %unicode support

@@

A \textit{chatbot} is a program, sometimes developed using Artificial Intelligence techniques, able to communicate in a similar way as humans do \cite{Gong2008}, using, in some cases, natural language. Indeed, most of these applications would be, apparently, close to passing a classic Turing test \cite{moor2003turing}, since they are able to answer almost any question fluently, and even pose own questions in their conversations. The recent advancement in this discipline has played an important role in many fields especially in education and online tutoring \cite{clarizia2018,Smutny2020}. These automatic systems facilitate the delivery of personalized learning by adapting to students’ pace of learning and providing customized online tutoring outside the classroom. Thus, chatbots can significantly contribute to providing interactive learning experiences as well as improving individual attention \cite{agarwal2020}.
 In this sense, chatbot technology offers a great opportunity for the improvement of tutoring systems \cite{Daniel2016,agarwal2020}, as not all students are comfortable with face-to-face tutoring with the instructor. In fact, in many cases, some students experience stress because of their need to ask a question in front of the entire class. This is why many students resort to silence  
 % Antonio - I guess refrain to silence is the opposite of what we wanted to say, isn't it? :D
 % mistaking "resort" for "refrain". You don't say "refrain to", but refrain from.
 in the classroom and tend to contact their teachers later via email.
 Consequently, this can lead to not only a delay in obtaining answers to their questions butcan also cause a significant increase in their teachers’ workload, especially when factoring in the disproportion between the class size and the number of teachers. Hence, the chatbot technology has a potential to mitigate this problem by providing answers to students’ questions and facilitating a dynamic and autonomous learning experience \cite{Griol2014,kim2020}. In addition, using an automated system such as a chatbot, can draw teachers’ attention to topics that students struggle to comprehend or need further assistance in understanding.

@@

% Describe the survey and population
We collected responses from 282 teachers from Spain and Spanish Speaking countries for the first survey (\textit{survey1} in Fig. \ref{figure:teachers_population}) and 205 for the second one (\textit{survey2}). The sample includes mostly teachers from university, although about 32\% and 8\% are from vocational education in the first and second survey respectively. Regarding gender, 61\% of the teachers were male and 32\% female (approximately 5\% chose not to disclose it). Finally, the responses are equally distributed according to the teaching experience, showing 24\% of responses from teachers with 5 or less years of experience, 26\% for 6-15 years of experience, 30\% for 16-25, and 20\% for teachers with more than 25 years of experience.

@@

% Fran: I think figures with data about the use of messaging apps should come here. Then, we say we also asked about changes after covid and include data about that
Results in Fig. \ref{figure:chatbots_use} do not show significant differences in the use of instant messaging apps between teachers from University and vocational education. In general, most teachers prefer messaging apps provided by their own institutions. With respect to specific disciplines, Engineering and Technology teachers are more active in their use, although it is also relevant the amount of teachers from Humanities that answered they used these apps in their classes (around 60\% use the apps provided by their institutions).

@@

    \begin{minipage}[t]{0.31\textwidth}

@@

ggplot(df1, aes(x = MessagingPlatformProvided, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male','PNTS')))+ theme(axis.title.y=element_blank())

@@

ggplot(df1, aes(x = MessagingPlatformProvided, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Experience)+ theme(axis.title.y=element_blank())

@@

    \begin{minipage}[t]{0.31\textwidth}

@@

<<sector_provided, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformProvided, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank())

@@

    %----
    \vspace{0.2mm}
    \hspace{10mm}
    \begin{minipage}[t]{0.5\textwidth}

@@

ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformProvided, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering", "Other")))+ theme(axis.title.y=element_blank())

@@

    \hspace{10mm}
    %----
    %----
    \vspace{0.5mm}
    \begin{minipage}[t]{0.31\textwidth}

@@

ggplot(df1, aes(x = MessagingPlatformWhatsApp, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male','PNTS')))+ theme(axis.title.y=element_blank())

@@

ggplot(df1, aes(x = MessagingPlatformWhatsApp, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Experience)+ theme(axis.title.y=element_blank())

@@

<<sector_whatsapp, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformWhatsApp, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank())

@@

    \vspace{0.2mm}
    \hspace{10mm}
    \begin{minipage}[t]{0.5\textwidth}

@@

ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformWhatsApp, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering", "Other")))+ theme(axis.title.y=element_blank())

@@

    \hspace{10mm}
    %----
    %----
    \vspace{0.5mm}
    \begin{minipage}[t]{0.31\textwidth}

@@

ggplot(df1, aes(x = MessagingPlatformTelegram, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male','PNTS')))+ theme(axis.title.y=element_blank())

@@

ggplot(df1, aes(x = MessagingPlatformTelegram, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Experience)+ theme(axis.title.y=element_blank())

@@

<<sector_telegram, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformTelegram, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank())

@@

ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformTelegram, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering", "Other")))+ theme(axis.title.y=element_blank())

@@

    \hspace{10mm}

@@

    \caption{Use of chatbots in class: distributions per gender, age, years of experience in education, university or vocational education, and discipline. Teachers were specifically asked about two popular instant messaging apps (Whatsapp and Telegram), and the use of platforms already provided by their institutions different than these two ones.}
\label{figure:chatbots_use}

@@

% This is about changes (due to) after covid
Some of the questions in the survey were focused on the impact of the COVID19 pandemic between 2020 and 2021, in the teachers' viewpoints towards the use of instant messaging apps int their class. The 282 
% Fran, please check Olia's comment as having 282 answers 
% -> I did remove some responses that Olia considered 'No' by default (i guess). I' ll keep it to 282 to be consistent
answers are summarized in Fig. \ref{figure:chatbots_use_postcovid}, showing that about 77\% of teachers already used these tools before the pandemic and kept using them during the pandemic lockdowns that forced students and educators to use remote education schemes. Moreover, approximately 15\% of them switched their messaging app for one that offered a safer interaction with their students. According to the responses, an additional 16\% started using messaging apps during the pandemic for the first time in their classes. Aggregating the answers, currently 91\% of the teachers are using messaging apps as learning tools. 
%A few relevant facts from the collected data: teachers with 16-25 years of experience seem more worried about the security of messaging apps and teachers with up to 5 years of experience were more open to adopt messaging apps due to the pandemic; female teachers were more willing to adopt new platforms due to the pandemic, and even to start using safer alternatives; finally, university teachers that did not use them were more reluctant to start using them after the pandemic.
\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.85\textwidth}
        \centering
<<gender_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
survey.pilot.postCOVID <- within(survey.pilot.postCOVID, PostCOVID.changes <- factor(PostCOVID.changes, levels=names(sort(table(PostCOVID.changes), decreasing=TRUE))))
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Gender, fill = Gender ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_covid.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.85\textwidth}
        \centering
<<sector_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Sector, fill = Sector ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/sector_covid.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.85\textwidth}
        \centering
<<experience_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Experience, fill = Experience ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@       
%       \includegraphics[width=\textwidth]{figures/experience_covid.pdf}
    \end{minipage}
\end{center}
    \caption{Impact of COVID in the use of messaging apps in University and vocational education.}
    % Have different sub-captions so that we can refer to specific subfigures - Olia
\label{figure:chatbots_use_postcovid}
\end{figure}

@@

% Olia - TODO: check for repetitions in the following paragraph that you added. thank you.
Finally, we took a closer look at the post-covid changes and the variables gender, sector, and discipline for which the chi-square test is statistically significant as shown in Figures \ref{figure:gender*postcovid_chisq}, \ref{figure:sector*postcovid_chisq}, \ref{figure:discipline*postcovid_chisq} respectively. Each figure shows on the left, a graph that represents the Pearson's residuals of the chi-square test results and a table on the right part that shows the contribution of each cell to the test. In the Pearson's residual graphs, the responses were shortened for the sake of clarity according to: 1) \textit{Did/do} stands for \textit{Yes, I did use it before and do use it now}; 2) \textit{Didn't/do} for \textit{Yes, I didn’t use any messaging app before, but do use it now}; 3) \textit{safer} for \textit{Yes, I use a safer messaging app now}; 4) \textit{Didn't/Don't} for \textit{No, I didn’t use it before and don’t use it now}.
%
\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[c]{0.40\textwidth}
        \centering
<<gender_postcovid_chisq, echo=FALSE, fig.width=17/2.54, fig.height=10/2.54>>=
df <-select(survey.pilot.postCOVID, PostCOVID.changes, Gender, Sector, Experience, Discipline)
df[as.character(df$Gender)!="PNTS", ] -> df1
df1[as.character(df1$Gender)!="No", ] -> df1
df1[as.character(df1$Discipline)!="No", ] -> df1
df1[as.character(df1$Experience)!="No", ] -> df1
df1[as.character(df1$Sector)!="No", ] -> df1
survey.pilot.postCovidChisq<- droplevels(df1)

>>>>

%%% Load packages
\usepackage{amsthm,amsmath}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[utf8]{inputenc} %unicode support

@@

A \textit{chatbot} is a program, sometimes developed using Artificial Intelligence techniques, able to communicate in a similar way as humans do \cite{Gong2008}, using, in some cases, natural language. Indeed, most of these applications would be, apparently, close to passing a classic Turing test \cite{moor2003turing}, since they are able to answer almost any question fluently, and even pose own questions in their conversations. The recent advancement in this discipline has played an important role in many fields especially in education and online tutoring \cite{clarizia2018,Smutny2020}. These automatic systems facilitate the delivery of personalized learning by adapting to students’ pace of learning and providing customized online tutoring outside the classroom. Thus, chatbots can significantly contribute to providing interactive learning experiences as well as improving individual attention \cite{agarwal2020}.
 In this sense, chatbot technology offers a great opportunity for the improvement of tutoring systems \cite{Daniel2016,agarwal2020}, as not all students are comfortable with face-to-face tutoring with the instructor. In fact, in many cases, some students experience stress because of their need to ask a question in front of the entire class. This is why many students resort to silence  
 % Antonio - I guess refrain to silence is the opposite of what we wanted to say, isn't it? :D
 % mistaking "resort" for "refrain". You don't say "refrain to", but refrain from.
 in the classroom and tend to contact their teachers later via email.
 Consequently, this can lead to not only a delay in obtaining answers to their questions but can also cause a significant increase in their teachers’ workload, especially when factoring in the disproportion between the class size and the number of teachers. Hence, the chatbot technology has a potential to mitigate this problem by providing answers to students’ questions and facilitating a dynamic and autonomous learning experience \cite{Griol2014,kim2020}. In addition, using an automated system such as a chatbot, can draw teachers’ attention to topics that students struggle to comprehend or need further assistance in understanding.

@@

% Describe the survey and population
We collected responses from 282 teachers from Spain and Spanish Speaking countries for the first survey (\textit{survey1} %in Fig. \ref{figure:teachers_population}
) and 205 for the second one (\textit{survey2}). The sample includes mostly teachers from university, although about 32\% and 8\% are from vocational education in the first and second survey respectively. Regarding gender, 61\% of the teachers were male and 32\% female (approximately 5\% chose not to disclose it). Finally, the responses are equally distributed according to the teaching experience, showing 24\% of responses from teachers with 5 or less years of experience, 26\% for 6-15 years of experience, 30\% for 16-25, and 20\% for teachers with more than 25 years of experience.

@@

% Fran: I think figures with data about the use of messaging apps should come here. Then, we say we also asked about changes after covid and include data about that
Results in Figs. \ref{figure:provided*chatbots_use}, \ref{figure:whatsapp*chatbots_use}, \ref{figure:telegram*chatbots_use} do not show significant differences in the use of instant messaging apps between teachers from University and vocational education. In general, most teachers prefer messaging apps provided by their own institutions (we also show the results for the two most popular messaging app platforms: Whatsapp and Telegram). With respect to specific disciplines, Engineering and Technology teachers are more active in their use, although it is also relevant the amount of teachers from Humanities that answered they used these apps in their classes (around 60\% use the apps provided by their institutions).

@@

    \begin{minipage}[t]{0.35\textwidth}

@@

df1[as.character(df1$Gender)!="PNTS", ] -> df1
ggplot(df1, aes(x = MessagingPlatformProvided, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male')))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
#+ facet_grid(~factor(Gender, levels=c('Female','Male','PNTS')))+ theme(axis.title.y=element_blank())

@@

ggplot(df1, aes(x = MessagingPlatformProvided, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Experience)+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())

@@

    \begin{minipage}[t]{0.3\textwidth}

@@

<<sector_provided, echo=FALSE, fig.width=10/2.54, fig.height=8.25/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformProvided, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank())+ theme(axis.title.x=element_blank())

@@

    \begin{minipage}[t]{0.6\textwidth}

@@

survey.pilot.Discipline[as.character(survey.pilot.Discipline$Discipline)!="Other", ] -> survey.pilot.Discipline
ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformProvided, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + #facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering", "Other")))+ theme(axis.title.y=element_blank())
facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering")))+ theme(axis.title.y=element_blank())+ theme(axis.title.x=element_blank())

@@

    \caption{Use of chatbots provided by the teachers' academic institutions in class: distributions per gender, years of experience in education, university or vocational education, and discipline.}
    \label{figure:provided*chatbots_use}
    \begin{minipage}[t]{0.35\textwidth}

@@

df1[as.character(df1$Gender)!="PNTS", ] -> df1
ggplot(df1, aes(x = MessagingPlatformWhatsApp, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male')))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())

@@

ggplot(df1, aes(x = MessagingPlatformWhatsApp, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Experience)+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())

@@

<<sector_whatsapp, echo=FALSE, fig.width=10/2.54, fig.height=8.25/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformWhatsApp, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())

@@

    \begin{minipage}[t]{0.6\textwidth}

@@

ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformWhatsApp, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering")))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())

@@

    \caption{Use of Whatsapp in class: distributions per gender, years of experience in education, university or vocational education, and discipline.}
    \label{figure:whatsapp*chatbots_use}
    \begin{minipage}[t]{0.35\textwidth}

@@

df1[as.character(df1$Gender)!="PNTS", ] -> df1
ggplot(df1, aes(x = MessagingPlatformTelegram, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male')))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())

@@

ggplot(df1, aes(x = MessagingPlatformTelegram, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Experience)+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())

@@

<<sector_telegram, echo=FALSE, fig.width=10/2.54, fig.height=8.25/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformTelegram, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())

@@

ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformTelegram, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering")))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())

@@

@@

    \caption{Use of Telegram in class: distributions per gender, years of experience in education, university or vocational education, and discipline.}
    \label{figure:telegram*chatbots_use}

@@

% This is about changes (due to) after covid
A question in the survey was also focused on the impact of the COVID19 pandemic between 2020 and 2021, in the teachers' viewpoints towards the use of instant messaging apps int their class. The 282 
% Fran, please check Olia's comment as having 282 answers 
% -> I did remove some responses that Olia considered 'No' by default (i guess). I' ll keep it to 282 to be consistent
answers are summarized in Figs. \ref{figure:gender*chatbots_use_postcovid}, \ref{figure:sector*chatbots_use_postcovid}, \ref{figure:years*chatbots_use_postcovid}, \ref{figure:disc*chatbots_use_postcovid}, showing that about 77\% of teachers already used these tools before the pandemic and kept using them during the pandemic lockdowns that forced students and educators to use remote education schemes. Moreover, approximately 15\% of them switched their messaging app for one that offered a safer interaction with their students. According to the responses, an additional 16\% started using messaging apps during the pandemic for the first time in their classes. Aggregating the answers, currently 91\% of the teachers are using messaging apps as learning tools. 
%A few relevant facts from the collected data: teachers with 16-25 years of experience seem more worried about the security of messaging apps and teachers with up to 5 years of experience were more open to adopt messaging apps due to the pandemic; female teachers were more willing to adopt new platforms due to the pandemic, and even to start using safer alternatives; finally, university teachers that did not use them were more reluctant to start using them after the pandemic.
    \begin{minipage}[t]{0.80\textwidth}
<<gender_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
survey.pilot.postCOVID <- within(survey.pilot.postCOVID, PostCOVID.changes <- factor(PostCOVID.changes, levels=names(sort(table(PostCOVID.changes), decreasing=TRUE))))
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Gender, fill = Gender ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90), size=7) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
%       \includegraphics[width=\textwidth]{figures/gender_covid.pdf}
        \vspace{-1cm}
        \caption{Total count of responses for the use of messaging apps after the COVID-19 pandemic grouped by gender (PNTS stands for Prefer Not To Say).}
        \label{figure:gender*chatbots_use_postcovid}
    \begin{minipage}[t]{0.80\textwidth}
<<sector_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Sector, fill = Sector ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
%       \includegraphics[width=\textwidth]{figures/sector_covid.pdf}
        \vspace{-1cm}
        \caption{Total count of responses for the use of messaging apps after the COVID-19 pandemic grouped by sector.}
        \label{figure:sector*chatbots_use_postcovid}
    \begin{minipage}[t]{0.80\textwidth}
<<experience_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Experience, fill = Experience ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
%       \includegraphics[width=\textwidth]{figures/experience_covid.pdf}
        \vspace{-1cm}
        \caption{Total count of responses for the use of messaging apps after the COVID-19 pandemic grouped by years of experience in teaching.}
        \label{figure:years*chatbots_use_postcovid}
        \begin{minipage}[t]{0.80\textwidth}
<<discipline_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
survey.pilot.postCOVID %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Technology", "Tech/Engineering", as.character(Discipline)))->survey.pilot.postCOVID
survey.pilot.postCOVID[as.character(survey.pilot.postCOVID$Discipline)!="No", ] -> survey.pilot.postCOVID
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Discipline, fill = Discipline ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
%       \includegraphics[width=\textwidth]{figures/sector_covid.pdf}
        \vspace{-1cm}
        \caption{Total count of responses for the use of messaging apps after the COVID-19 pandemic grouped by discipline.}
        \label{figure:disc*chatbots_use_postcovid}
    % Have different sub-captions so that we can refer to specific subfigures - Olia

@@

% Olia - TODO: check for repetitions in the following paragraph that you added. thank you.
Finally, we took a closer look at the post-covid changes and the variables gender, sector, and discipline for which the chi-square test is statistically significant as shown in Figures \ref{figure:gender*postcovid_chisq}, \ref{figure:sector*postcovid_chisq}, \ref{figure:discipline*postcovid_chisq} respectively. Each figure shows on the left, a graph that represents the Pearson's residuals of the chi-square test results and a table on the right part that shows the contribution of each cell to the test. In the Pearson's residual graphs, the responses were shortened for the sake of clarity according to: 1) \textit{Did/do} stands for \textit{Yes, I did use it before and do use it now}; 2) \textit{Didn't/do} for \textit{Yes, I didn’t use any messaging app before, but do use it now}; 3) \textit{safer} for \textit{Yes, I use a safer messaging app now}; 4) \textit{Didn't/Don't} for \textit{No, I didn’t use it before and don’t use it now}.
%
\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[c]{0.45\textwidth}
        \centering
<<gender_postcovid_chisq, echo=FALSE, fig.width=17/2.54, fig.height=10/2.54>>=
df <-select(survey.pilot.postCOVID, PostCOVID.changes, Gender, Sector, Experience, Discipline)
df[as.character(df$Gender)!="PNTS", ] -> df1
df1[as.character(df1$Gender)!="No", ] -> df1
df1[as.character(df1$Discipline)!="No", ] -> df1
df1[as.character(df1$Experience)!="No", ] -> df1
df1[as.character(df1$Sector)!="No", ] -> df1
survey.pilot.postCovidChisq<- droplevels(df1)

covidChanges.vs.Gender<-table(survey.pilot.postCovidChisq$Gender, survey.pilot.postCovidChisq$PostCOVID.changes)
colnames(covidChanges.vs.Gender)<- c("Did/do", "Didn't/do","safer", "Didn't/don't")
covidChanges.vs.Gender<-covidChanges.vs.Gender[,c(3,2,1,4)]
chisq<-chisq.test(covidChanges.vs.Gender)
corrplot(chisq$residuals, is.corr = FALSE)
@
    \end{minipage}
    %This tables comes from (TODO: generate them automatically from the data)
    %contrib <- 100*chisq$residuals^2/chisq$statistic
    %round(contrib, 2)
    %round(chisq$stdres, 3)
    \begin{minipage}[c]{0.50\textwidth}
        \centering
    \resizebox{\textwidth}{!}{%
\begin{tabular}{llllll}
\hline
 &  & \multicolumn{4}{c}{Gender} \\ \cline{3-6} 
 &  & \multicolumn{2}{c}{Female} & \multicolumn{2}{c}{Male} \\ \cline{3-6} 
 &  & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Stand. \\ Residuals\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Contri-\\ bution\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Stand. \\ Residual\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Contri-\\ bution\end{tabular}} \\ \cline{2-6} 
\begin{tabular}[c]{@{}l@{}}Post-\\ covid\\ changes\end{tabular} & \begin{tabular}[c]{@{}l@{}}Yes, I use a \\ safer messaging \\ app now\end{tabular} & 1.832 & 13.49\% & -1.832 & 7.50\% \\
 & \begin{tabular}[c]{@{}l@{}}Yes, I didn’t use \\ any messaging \\ app before, but\\ do use it now\end{tabular} & \textbf{2.806} & \textbf{31.21\%} & \textbf{-2.806} & \textbf{17.36\%} \\
 & \begin{tabular}[c]{@{}l@{}}No, I did use it \\ before and\\ use it now\end{tabular} & \textbf{2.850} & 15.15\% & \textbf{-2.850} & 8.43\% \\
 & \begin{tabular}[c]{@{}l@{}}No, I didn’t use \\ it before and\\ don’t use it now\end{tabular} & -1.011 & 4.40\% & 1.011 & 2.45\% \\ \hline
\end{tabular}
    %\begin{tabular}{l|r|r|r|r|}
    %\cline{2-5}
    %                             & Did/Do & Didn't/Do & Safer  & Didn' t/Don't \\ \hline
    %\multicolumn{1}{|l|}{Female} & \textbf{-2.850} & \textbf{2.806} & 1.832  & -1.011 \\ 
    %\multicolumn{1}{|l|}{} & 15.15\% & \textbf{31.21\%} &  13.49\% & 4.40\% \\ \hline
    %\multicolumn{1}{|l|}{Male}   & \textbf{2.850} & \textbf{-2.806} & -1.832 & 1.011 \\ 
    %\multicolumn{1}{|l|}{} &  8.43\% &  \textbf{17.36\%} & 7.50\% &  2.45\% \\ \hline
    %\end{tabular}%
    }
    \end{minipage}
\end{center}
    \caption{Analysis for gender and post-covid changes: left) Pearson's residuals for the Chi-square test; right) Gender*post-covid changes crosstabulation, including standardized residuals \cite{agresti2013categorical} (also adjusted standardized residuals) and contribution percentage to the total Chi-square test of each cell.}
\label{figure:gender*postcovid_chisq}
\end{figure}
